---
title: "ce_cleaning_registration_data_all_fuels"
output: html_document
description: Read and clean data on vehicle registrations - new vehicle registrations including all fuel types.
---

Load libraries
```{r}
library(tidyverse)
library(dplyr)
library(plotly)
library(ggplot2)
library(httr)
#library(jsonlite)
library(stringr)
library(rvest)
library(xml2) 
```
Define function to cleanup column names of files exported from data provider
```{r}
#custom function to cleanup column names
colname_cleanup <- function(x) {
    colnames(x) <- x %>%
      colnames() %>%
      str_to_lower() %>% #make colnames lowercase
      str_replace_all(" ", "_") %>% #replace spaces with underscores
      str_replace_all("\\/", "_") %>% #replace slashes with underscores
      str_replace_all("calendar_year_\\d{4}_vehicle_count", "vehs_reg") %>%
      str_replace_all("calendar_ytd_\\d{4}_vehicle_count", "vehs_reg")
    ;x 
    }

```

List makes & models that should be excluded to use as filter
```{r}
exclude_make <- c("ADVANCE MIXER",
                  "AMERICAN LAFRANCE",
                  "AUTOCAR",
                  "BYD TRANSIT",
                  "GEM", 
                  "KALMAR", 
                  "LION BUS",
                  "MITSUBISHI FUSO",
                  "CAPACITY OF TEXAS", 
                  "CATERPILLAR",
                  "CRANE CARRIER",
                  "ELDORADO",
                  "EMERGENCY ONE",
                  "FERRARA",
                  "FREIGHTLINER",
                  "GILLIG",
                  "HENDRICKSON",
                  "HINO",
                  "IC CORP",
                  "INTERNATIONAL",
                  "ISUZU",
                  "KENWORTH",
                  "KIMBLE",
                  "KOVATCH",
                  "MACK",
                  "NISSAN UD",
                  "OSHKOSH",
                  "PETERBILT",
                  "PIERCE",
                  "ROADMASTER RAIL",
                  "ROSENBAUER",
                  "SPARTAN MOTORS",
                  "SUZUKI",
                  "TEMSA",
                  "TIFFIN",
                  "THOMAS BUILT BUSES",
                  "WESTERN STAR",
                  "WORKHORSE")

exclude_model <- c("LCF","F53 MOTORHOME", "F59", "BUS", "VHD", "VN", "VNR")
```

Read in files with county-level vehicle registrations by fuel type for 2015-2020
```{r}
co_mmy_fuel_19 <- read_csv('../data/county_mmy_fuel_19.csv')%>%
  colname_cleanup()%>%
  mutate(reg_year="2019")%>%
  mutate(reg_year = reg_year%>%as.numeric())
glimpse(co_mmy_fuel_19)

co_mmy_fuel_20 <- read_csv('../data/county_mmy_fuel_20.csv')%>%
  colname_cleanup()%>%
  mutate(reg_year="2020")%>%
  mutate(reg_year = reg_year%>%as.numeric())
glimpse(co_mmy_fuel_20)

co_mmy_fuel_18 <- read_csv('../data/county_mmy_fuel_18.csv')%>%
  colname_cleanup()%>%
  mutate(reg_year="2018")%>%
  mutate(reg_year = reg_year%>%as.numeric())
glimpse(co_mmy_fuel_18)

co_mmy_fuel_17 <- read_csv('../data/county_mmy_fuel_17.csv')%>%
  colname_cleanup()%>%
  mutate(reg_year="2017")%>%
  mutate(reg_year = reg_year%>%as.numeric())
glimpse(co_mmy_fuel_17)

co_mmy_fuel_16 <- read_csv('../data/county_mmy_fuel_16.csv')%>%
  colname_cleanup()%>%
  mutate(reg_year="2016")%>%
  mutate(reg_year = reg_year%>%as.numeric())
glimpse(co_mmy_fuel_16)

co_mmy_fuel_15 <- read_csv('../data/county_mmy_fuel_15.csv')%>%
  colname_cleanup()%>%
  mutate(reg_year="2015")%>%
  mutate(reg_year = reg_year%>%as.numeric())
glimpse(co_mmy_fuel_15)

#Combine each year's data
co_mmy_fuel <- bind_rows(co_mmy_fuel_20, 
                         co_mmy_fuel_19, 
                         co_mmy_fuel_18, 
                         co_mmy_fuel_17, 
                         co_mmy_fuel_16, 
                         co_mmy_fuel_15)%>%
                filter(make %in% exclude_make == FALSE) #Exclude irrelevant brands

#Adjust fuel names
co_mmy_fuel$fuel <- co_mmy_fuel$fuel%>%
  str_replace_all("FLEXIBLE \\(GASOLINE\\/ETHANOL\\)", "FLEX FUEL")%>%
  str_replace_all("BOTH GAS AND ELECTRIC", "HYBRID")

glimpse(co_mmy_fuel)
```

County names do not come with FIPS codes. Create geoid key and map against county labels for plotting.
```{r}
#Map FIPS codes to county labels using an existing dataset.

county_fips <- read_csv('../data/county_tracts.csv') %>%
  colname_cleanup() %>% #apply colname cleanup function
  select(!vehicle_count) %>% #exclude vehicle count column
  mutate(countyid = census_tracts%>%str_extract("^\\d{5}")) %>% #extract FIPS from tract id
  select(county, countyid)%>% 
  distinct(.keep_all = False) #select unique records --> 3,220 county geoids
 
glimpse(county_fips)
```

Join co_mmy_fuel with county_fips
```{r}
#Join dataframes.

co_mmy_fuel <- left_join(co_mmy_fuel, county_fips)%>%
  select(county, countyid, make, model, model_year, fuel, vehs_reg, reg_year)

#Parse out state identifiers
co_mmy_fuel <- co_mmy_fuel %>% mutate(state = county%>%str_extract("^([A-Z]{2})*"), #extract state abbreviation
                                      stateid = countyid %>% str_extract("^\\d{2}") #extract state code
                                      ) %>%
                                filter(state!="PR") %>% #exclude Puerto Rico
  select(state, stateid, county, countyid, make, model, model_year, fuel, vehs_reg, reg_year)

glimpse(co_mmy_fuel)
```

Write completed co_mmy_fuel dataframe to CSV
```{r}
#Write to CSV
write.csv(co_mmy_fuel, '../data/co_mmy_fuel_df.csv', row.names = FALSE)
```


#Prepare for use in webscraping MSRP data.
Deduplicate make/model/year/fuel combinations
```{r}
dedupe_mmy <- function(x) {
  x <- x %>%
    filter(make %in% exclude_make == FALSE) %>%
    filter(model %in% exclude_model ==FALSE) %>%
    select(make, model, fuel, model_year) %>%
    distinct(.keep_all = False) %>%
    group_by(make) %>%
    arrange(make, model, model_year, .by_group=TRUE)%>%
    rename(ihs_make = make, ihs_model = model)
  ;x
}

non_ev_models <- co_mmy_fuel%>%dedupe_mmy()%>%  #5.4M rows down to 3531 rows
  filter(fuel!="ELECTRIC") # down to 3,418 rows

view(non_ev_models)
glimpse(non_ev_models)
```

Scroll through makes & models to check for discrepancies with Autotrader formatting
```{r}
#Check makes
makes <- unique(non_ev_models$ihs_make)
length(makes)  #44 makes
view(makes)

#check models
models<- unique(non_ev_models$ihs_model)
length(models) #487 models
view(models)

non_ev_models%>%group_by(ihs_make)%>%tally()%>%ungroup()

non_ev_models%>%
  select(ihs_make, ihs_model, model_year)%>%
  group_by(ihs_make)%>%
  tally()%>%
  ungroup()
```



Copy make & model columns to match Autotrader string formatting & generate URL.
```{r}

non_ev_models<-non_ev_models %>%
  #create columns for autotrader make and model, cast to lowercase & normalize formatting
  mutate(autdr_make = 
           str_to_lower(ihs_make)%>% 
           str_replace("smartcar","smart")%>% #smart instead of smartcar
           str_replace(" ","-"), #alfa romeo -> alfa-romeo
         autdr_model=
           str_to_lower(ihs_model)%>%
           str_remove(" ev")%>% #kia soul vs kia soul ev
           str_remove("mini ")%>% #cooper vs mini cooper
           str_remove(" denali")%>%
           str_replace(" ","-")%>% #model-3 vs model 3, #mustang-mach-e
           str_replace("a4-allroad", "allroad")%>%
           str_replace("q5-e", "q5")%>%
           str_replace("trail-blazer", "trailblazer")%>%
           str_replace("town-& country", "town-&-country")%>%
           str_replace("458", "458-italia")%>%
           str_replace("acadia-limited", "acadia")%>%
           str_replace("xfr", "xf")%>%
           str_replace("ct200h", "ct-200h")%>%
           str_replace("4-runner", "4runner")%>%
           str_replace("mp4", "mp4-12c")%>%
           str_replace("frontier-/xe", "frontier")%>%
           str_remove("-plug-in|-ia|-cross country|-sportwagen|-limited|-select|-coupe|-gt|-sedan| sport| xl|-n")%>%
           str_replace("clubman", "cooper-clubman")%>%
           str_replace("paceman", "cooper-paceman")%>%
           str_replace("countryman", "cooper-countryman")%>%
           str_replace("roadster", "cooper-roadster")%>%
           str_replace("coupe", "cooper-coupe")%>%
           str_replace("^2$", "mazda2")%>%
           str_replace("^3$", "mazda3")%>%
           str_replace("^5$", "mazda5")%>%
           str_replace("^6$", "mazda6")%>%
           str_replace("^mx5$", "mx-5-miata")
           )%>%
  #add column, constructing vehicle url on autotrader.com
  mutate(
    autotrader_url=paste(
      'https://www.autotrader.com', autdr_make, autdr_model, model_year, sep = '/')
    )


glimpse(non_ev_models)
view(non_ev_models)

```

#EXTRACT ROWS FOR ROUND 1 SCRAPING
```{r}
problem_makes <- c("bmw", "ferrari", "land rover")
problem_models <- c("savana", "sierra", "f-series")

filter(make %in% exclude_make == FALSE) %>%
    filter(model %in% exclude_model ==FALSE)
  
round1 <- non_ev_models%>%
  filter(make%in% problem_makes == FALSE) %>%
  ilter(model %in% problem_models ==FALSE)

hybrids <- non_ev_models %>%
  filter(fuel=="HYBRID")

glimpse(hybrids)
hybrids
```


#CLEAN AND FORMAT MIDDLE TN DETAILED REGISTRATION
```{r}
mdtn_mmyf_19 <- read_csv('../data/mdtn_mmy_fuel_19.csv')%>%
  colname_cleanup()%>%
  filter(make %in% exclude_make == FALSE)%>% #Exclude irrelevant makes
  filter(make %in% exclude_model == FALSE)%>% #Exclude irrelevant models
  rename(geoid=census_tracts)%>% #Rename tracts to geoid in order to join with shapefiles
  rename(vehs_reg = calendar_year_2019_vehicle_count)

#Adjust fuel names
mdtn_mmyf_19$fuel <- mdtn_mmyf_19$fuel%>%
  str_replace_all("FLEXIBLE \\(GASOLINE\\/ETHANOL\\)", "FLEX FUEL")%>%
  str_replace_all("BOTH GAS AND ELECTRIC", "HYBRID")

glimpse(mdtn_mmyf_19)
unique(mdtn_mmyf_19$fuel)
```

The resulting dataframe includes new vehicle registrations for all fuel types in the 37 counties of Middle TN

```{r}

length(mdtn_mmyf_19$county%>% unique())

length(mdtn_mmyf_19$fuel%>% unique())
```

```{r}
write.csv(mdtn_mmyf_19, '../data/mdtn_mmyf_19_df.csv', row.names = FALSE)
```

