---
title: "ce_cleaning_registration_data"
output: html_document
description: Read and clean data on vehicle registrations
---

Load libraries
```{r}
library(tidyverse)
library(dplyr)
library(plotly)
library(ggplot2)
library(httr)
#library(jsonlite)
library(stringr)
```

Define custom function to cleanup column names of files exported from data provider
```{r}
#custom function to cleanup column names
colname_cleanup <- function(x) {
    colnames(x) <- x %>%
      colnames() %>%
      str_to_lower() %>% #make colnames lowercase
      str_replace_all(" ", "_") %>% #replace spaces with underscores
      str_replace_all("\\/", "_") #replace slashes with underscores
    ;x 
    }

```


Read in CSV file with EV registrations for TN & GA Jan 2015-Oct 2020
```{r}
tnga_ev_reg <- read_csv('../data/ev_tn_ga_2015_2020.csv')
glimpse(tnga_ev_reg)

tnga_ev_reg<-colname_cleanup(tnga_ev_reg)
glimpse(tnga_ev_reg)

```

Read in CSV file with nationwide EV registrations Jan 2015-Oct 2020
```{r}
#remove spaces, slashes and uppercase characters
us_ev_reg <- read_csv('../data/retail_ev_us_15_20.csv')%>%
  colname_cleanup()

#List makes that should be excluded to use as filter
exclude_make <- c("BYD TRANSIT","GEM", "KALMAR", "LION BUS","MITSUBISHI FUSO")
us_ev_reg <- us_ev_reg%>%
  filter(make %in% exclude_make == FALSE)

#shorten column names
colnames(us_ev_reg) <- colnames(us_ev_reg)%>%
  str_replace_all("vehicle_count_calendar_year", "reg")%>%
  str_replace_all("vehicle_count_calendar", "reg")%>%
  str_replace("census_tracts","geoid")

glimpse(us_ev_reg)

#Write to CSV
write.csv(us_ev_reg, '../data/us_ev_reg_df.csv', row.names = FALSE)
```

Read in CSV file with (aggregated) nationwide registrations by fuel type Jan 2015-Nov 2020
```{r}
#remove spaces, slashes and uppercase characters
fueltype_reg <- read_csv('../data/fueltype_reg_2015_2020.csv')%>%
  colname_cleanup()

#shorten column names
colnames(fueltype_reg) <- colnames(fueltype_reg)%>%
  str_replace_all("_vehicle_count", "")%>%
  str_replace_all("calendar_year", "reg")%>%
  str_replace_all("calendar_ytd", "reg")%>%
  str_replace("census_tracts","geoid")

glimpse(fueltype_reg)
fueltype_reg

#Write to CSV
write.csv(fueltype_reg, '../data/fueltype_reg_sum_df.csv', row.names = FALSE)

```

Read in CSV file with (aggregated) nationwide mmy registrations by fuel type Jan 2015-Nov 2020
```{r}
#remove spaces, slashes and uppercase characters
mmy_fuel_reg <- read_csv('../data/mmy_fuel_reg_15_20.csv')%>%
  colname_cleanup()

#List makes that should be excluded to use as filter
exclude_make <- c("ADVANCE MIXER",
                  "AMERICAN LAFRANCE",
                  "AUTOCAR",
                  "BYD TRANSIT",
                  "GEM", 
                  "KALMAR", 
                  "LION BUS",
                  "MITSUBISHI FUSO",
                  "CAPACITY OF TEXAS", 
                  "CATERPILLAR",
                  "CRANE CARRIER",
                  "ELDORADO",
                  "EMERGENCY ONE",
                  "FERRARA",
                  "FREIGHTLINER",
                  "GILLIG",
                  "HENDRICKSON",
                  "HINO",
                  "IC CORP",
                  "INTERNATIONAL",
                  "ISUZU",
                  "KENWORTH",
                  "KIMBLE",
                  "KOVATCH",
                  "MACK",
                  "NISSAN UD",
                  "OSHKOSH",
                  "PETERBILT",
                  "PIERCE",
                  "ROADMASTER RAIL",
                  "ROSENBAUER",
                  "SPARTAN MOTORS",
                  "SUZUKI",
                  "TEMSA",
                  "THOMAS BUILT BUSES",
                  "WESTERN STAR",
                  "WORKHORSE")

mmy_fuel_reg <- mmy_fuel_reg%>%
  filter(make %in% exclude_make == FALSE)

#shorten column names
colnames(mmy_fuel_reg) <- colnames(mmy_fuel_reg)%>%
  str_replace_all("_vehicle_count", "")%>%
  str_replace_all("calendar_year", "reg")%>%
  str_replace_all("calendar_ytd", "reg")

glimpse(mmy_fuel_reg)
head(mmy_fuel_reg)
#dedupe_mmy(mmy_fuel_reg)%>%
 # select(ihs_make)%>%
  #distinct(.keep_all = False)

#Write to CSV
write.csv(mmy_fuel_reg, '../data/mmy_fuel_reg.csv', row.names = FALSE)
```


#Read in 2019 registrations of gasoline vehicles for TN
```{r}
tn_gas_reg_19 <- read_csv('../data/gas_reg_tn_2019.csv')%>%
  colname_cleanup()%>%
  mutate(fuel="gasoline")

colnames(tn_gas_reg_19) <- colnames(tn_gas_reg_19)%>%
  str_replace_all("calendar_year", "reg")%>%
  str_replace_all("_vehicle_count", "")%>%
  str_replace("census_tracts","geoid")

glimpse(tn_gas_reg_19)
write.csv(tn_gas_reg_19, '../data/tn_gas_reg_19_df.csv', row.names = FALSE)
```

Define function to de-dupe make/model/year entries in IHS registration data 

```{r}

#List makes that should be excluded to use as filter
exclude_make <- c("BYD TRANSIT","GEM", "KALMAR", "LION BUS","MITSUBISHI FUSO")

#(f) for distinct make/model/year values

dedupe_mmy <- function(x) {
  x <- x %>%
    filter(make %in% exclude_make == FALSE) %>%
    select(make, model, model_year) %>%
    distinct(.keep_all = False) %>%
    group_by(make) %>%
    arrange(make, model, model_year, .by_group=TRUE)%>%
    rename(ihs_make = make, ihs_model = model)
  ;x
  }

```

Apply de-dupe function to US EV registration data
```{r}

#US registrations
us_ev_mmy <- dedupe_mmy(us_ev_reg)

glimpse(us_ev_mmy)
view(us_ev_mmy)

#Tn & GA registrations - smaller data set to use for testing if necessary
#tnga_ev_models<- dedupe_mmy(tnga_ev_reg)
#View(tnga_ev_models)
#glimpse(tnga_ev_models)
```

Check make and model column for formatting differences with Autotrader data. 
```{r}
#Dataframe of makes
ev_ihs_makes <- unique(us_ev_mmy$ihs_make)
length(ev_ihs_makes)
view(ev_ihs_makes)

#Dataframe of models
ev_ihs_models <- unique(us_ev_mmy$ihs_model)
length(ev_ihs_models)
view(ev_ihs_models)

```

Copy make & model columns to match Autotrader string formatting & generate URL.
```{r}

us_ev_mmy<-us_ev_mmy %>%
  #create columns for autotrader make and model, cast to lowercase & normalize formatting
  mutate(autdr_make = 
           str_to_lower(ihs_make)%>% 
           str_replace("smartcar","smart"), #smart instead of smartcar
         autdr_model=
           str_to_lower(ihs_model)%>%
           str_remove(" ev")%>% #kia soul vs kia soul ev
           str_remove("mini ")%>% #cooper vs mini cooper
           str_replace(" ","-") #model-3 vs model 3, #mustang-mach-e 
           )%>%
  #add column, constructing vehicle url on autotrader.com
  mutate(
    autotrader_url=paste(
      'https://www.autotrader.com', autdr_make, autdr_model, model_year, sep = '/')
    )

view(us_ev_mmy) 
```

Write to CSV
```{r}
write.csv(us_ev_mmy, '../data/us_ev_mmy.csv', row.names = FALSE)
```


#LINKS & RESOURCES

https://stackoverflow.com/questions/11369961/subset-with-unique-cases-based-on-multiple-columns
https://stackoverflow.com/questions/7790732/unique-for-more-than-one-variable
https://stackoverflow.com/questions/34444295/how-to-specify-does-not-contain-in-dplyr-filter

#SCRAPS - unused/sample code

How many census tracts in US have EV registrations?
```{r}
length(us_ev_reg$census_tracts %>% unique())
```
How many Honda Clarity registrations are there in this set?
```{r}
us_ev_reg %>%
  select(!state)%>%
  select(!county)%>%
  select(!census_tracts)%>%
  select(!model_year)%>%
  group_by(model)%>%
  filter(model=="CLARITY")%>%
  summarise_if(is.numeric, sum, na.rm=TRUE)%>%
  mutate(total_registrations=rowSums(select(., starts_with("vehicle"))))
    
```

Read in file with sample vehicle data
```{r}
sample_cars<- read_csv('../data/test_vehicle_data.csv')%>%select(Make, Model,'Model Year')
view(sample_cars)
glimpse(sample_cars)
```

```{r}
#Original function for replacing space with underscore in column names
spaceless <- function(x) {colnames(x) <- gsub(" ", "_", colnames(x));x}

#Column cleanup function - alternate w/slight variation in first argument
colname_cleanup <- function(x) {
  colnames(x) <- str_to_lower(colnames(x)) %>% 
    str_replace_all(" ", "_") %>%
    str_replace_all("\\/", "_")
  ;x}



#ev_reg[!duplicated(ev_reg[4:6]),]%>%select(Make, Model,'Model Year')

#filter(location%in%input$location)%>%
```

