---
title: "ce_read_hhi_data"
output: html_document
date: 1/7/21
---

## Read in household income data by Census tract from ACS5 2019

```{r cars}
library(tidycensus)
library(tidyverse)
library(dplyr)
library(plotly)
library(ggplot2)
library(readxl)
library(leaflet)
library(sf)
```

```{r}
# Check your API key
Sys.getenv("CENSUS_API_KEY")
```
Pull in ACS5 2019 Subject Table variables
```{r}
var_19_acs5_sub <- load_variables(year=2019, dataset="acs5/subject",cache=TRUE)
View(var_19_acs5_sub)

```

Tidycensus has a built in table of counties/states and FIPS codes.
```{r}

tidycensus::fips_codes

view(fips_codes%>%
    select(state, state_code) %>%
    distinct(.keep_all = False))

state_code_list <- pull(fips_codes %>%
                          select(state_code)%>%
                          distinct(.keep_all = FALSE), #dedupe state_code column
                        state_code
                        )[1:51] #exclude US territories

```
Read in selected variables ACS5 2019 Subject Table S1901
```{r}
#s1901_vars <- c(total_households="S1901_C01_001", median_hhi = "S1901_C01_012",mean_hhi="S1901_C01_013")
#acs5_19_s1901

census_hhi <-get_acs(survey = "acs5",
                        year=2019, 
                        #table="S1901",
                        variables=c(total_households="S1901_C01_001",
                                     median_hhi = "S1901_C01_012",
                                     mean_hhi="S1901_C01_013"),
                        geography="tract", 
                        state=state_code_list,
                        cache=TRUE)
                        #show_call=TRUE
                        #output="wide"

#use colname_cleanup function to standardize column names and make lowercase
census_hhi <- colname_cleanup(census_hhi)

glimpse(census_hhi)

write.csv(census_hhi, '../data/census_hhi_df.csv', row.names = FALSE)
```


```{r}
tn_hhi_g <-get_acs(survey = "acs5",
                        year=2019, 
                        #table="S1901",
                        variables=c(total_households="S1901_C01_001",
                                     median_hhi = "S1901_C01_012",
                                     mean_hhi="S1901_C01_013"),
                        geography="tract", 
                        state="TN",
                        cache=TRUE,
                        geometry = TRUE)%>%
                  colname_cleanup()
head(tn_hhi_g)
glimpse(tn_hhi_g)
class(tn_hhi_g) #sf data.frame
summary(tn_hhi_g)

#st_write(tn_hhi_g, paste0(tempdir(), "../data", "nc.shp"))
st_write(tn_hhi_g, '../data/tn_hhi_g.shp', row.names = FALSE)

test_read <- st_read("../data/tn_hhi_g.shp")
glimpse(test_read)
class(test_read)
```
#TOO BIG TO READ IN ALL AT ONCE
```{r}
#census_hhi_g <-get_acs(survey = "acs5",
                       # year=2019, 
                        #table="S1901",
                       # variables=c(total_households="S1901_C01_001",
                                    # median_hhi = "S1901_C01_012",
                                    # mean_hhi="S1901_C01_013"),
                       #geography="tract",
                      # state=state_code_list,
                       #geometry=TRUE,
                      # cache=TRUE)

#glimpse(census_hhi_g)
#view(census_hhi_g)
```

Test plotting Davidson County with ggplotly
```{r}
p<-tn_hhi_g %>%
  filter(variable=="median_hhi") %>%
  filter(str_detect(name,"Davidson County")) %>%
  ggplot(aes(fill=estimate)) +
  geom_sf(color=NA) +
  #coord_sf(crs=)
  scale_fill_viridis_c(option="plasma",
                       labels=scales::dollar)+
  theme_minimal() +
  labs(title="Tract-Level Median HHI",
       fill= "Median HHI")

fig <- ggplotly(p) 
fig
```
Test plotting Davidson County with Leaflet.
```{r}

tn_pal <- colorNumeric(palette = "viridis",
                       domain=tn_hhi_g$estimate)

tn_hhi_g%>%
  filter(variable=="median_hhi") %>%
  leaflet()%>%
  addTiles()%>%
  addPolygons(stroke=FALSE,
              fillOpacity=0.8,
              smoothFactor = 0.5,
              color= ~tn_pal(estimate),
              label = ~paste0("Median HHI: ", estimate),
              highlight = highlightOptions(weight=3,
                                           color="red",
                                           bringToFront = TRUE))%>%
  setView(lng= -86.7833322,
          lat=36.1500571,
          zoom=9)

dollar()
```



Test plotting State of TN
```{r}
p <- s1901g_tn %>%
  filter(variable=="median_hhi") %>%
  #filter(str_detect(NAME,"Davidson County")) %>%
  ggplot(aes(fill=estimate)) +
  geom_sf(color=NA) +
  #coord_sf(crs=)
  scale_fill_viridis_c(option="plasma",
                       labels=scales::dollar)+
  theme_minimal() +
  labs(title="Tract-Level Median HHI",
       fill= "Median HHI")

fig <- ggplotly(p)
fig
```

Read in TN tract shapefile
```{r}
tn_tract_shape <- st_read("../data/tl_2019_47_tract/tl_2019_47_tract.shp")%>%
  colname_cleanup()
glimpse(tn_tract_shape)
class(tn_tract_shape)
#view(tn_tract_shape)
```


#Links & Sources

<https://mattherman.info/blog/tidycensus-mult/>
< https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html >
< https://github.com/walkerke/tidycensus/issues/72 > 
< https://sonoshah.github.io/tidycensuspresentation/slides#53 > 